Statement of Intent

### 2010-12-13 applications/phd_programs/berkeley_personal_history.tex
A lot of work on applications

My time working in the Applied Machine Learning lab has been the defining
experience of my undergraduate education.  I had very close interaction with
several professors and many graduate students who showed me how interesting
research can be.  From those that I had association with I gained a love of
learning new things and exploring new ideas.  I also learned a great deal more
than was taught in my classes in order to understand the research that was
going on in the lab.  Several times in my classes the professor teaching the
class had me give part of a lecture that dealt with things I was actively
researching.

>>>>

My time working in the Applied Machine Learning lab was the defining experience
of my undergraduate education.  I had very close interaction with several
professors and many graduate students who showed me how interesting research
can be.  From those that I had association with I gained a love of learning new
things and exploring new ideas.  I also learned a great deal more than was
taught in my classes in order to understand the research that was going on in
the lab.  Several times in my classes the professor teaching the class had me
give part of a lecture that dealt with topics I was actively researching.

### 2010-12-13 applications/phd_programs/personal_statement.tex
A lot of work on applications

I had my first major setback when that paper was rejected outright.  At this
point I had spent two years working on PSO, and I felt like giving up and
moving on.  PSO wasn't my main interest anyway, and I wanted to do more
interesting things.  My advisor, however, insisted that I perservere, telling
me that the experience would be good practice for a PhD.  So I revised my paper
and submitted it to another top conference, where it was
accepted~\cite{gardner-2010-speculative-evaluation-in-pso}.

@@

My work in particle swarm optimization was not directly applicable to this
program to which I am currently applying, though a general knowledge of
optimization is certainly useful in many fields of computer science.  However,
I feel that the experience I had working on that problem has done a great deal
to prepare me to succeed in your program.  I have experience conducting real
research, I know how to bring ideas to fruition, and I have successfully taken
my own innovations and had them recognized by the wider research community.
I am confident that, while there will inevitably be setbacks in any research
endeavor, I can perservere and succeed.

\bibliographystyle{plain}
\bibliography{bib}

>>>>

I had my first major setback when that paper was rejected outright.  At this
point I had spent two years working on PSO, and I felt like giving up and
moving on.  PSO wasn't my main interest anyway, and I wanted to do more
interesting things.  My advisor, however, insisted that I persevere, telling
me that the experience would be good practice for a PhD.  So I revised my paper
and submitted it to another top conference, where it was
accepted~\cite{gardner-2010-speculative-evaluation-in-pso}.

@@

My work in particle swarm optimization was not directly applicable to this
program to which I am currently applying, though a general knowledge of
optimization techniques is certainly useful in many fields of computer science.
However, I feel that the experience I had working on that problem has done a
great deal to prepare me to succeed in your program.  I have experience
conducting real research, I know how to bring ideas to fruition, and I have
successfully taken my own innovations and had them recognized by the wider
research community.  I am confident that, while there will inevitably be
setbacks in any research endeavor, I can persevere and succeed.

I have had many other experiences that also contribute to my preparation for
your program.  I have gained a broad background in statistical machine learning
techniques, having been the TA for the Bayesian Inference class taught at BYU
and taken several advanced classes on the subject.  I have also taken all of
the classes that BYU offers in natural language processing (NLP).  For my final
project in the graduate-level NLP course, I implemented a Gibbs sampler for the
latent Dirichlet allocation model and ran it on a novel dataset of religious
discourses.  That work was of particular interest to BYU, and I presented it at
the BYU Studies Symposium, an international conference held at
BYU~\cite{gardner-2010-general-conference-topics}.

As a masters student at BYU I have continued my work with topic modeling,
exploring how topic models can be used to improve the browsing experience of
large corpora.  I recently returned from the NIPS conference, where I presented
at a workshop a tool that I and others at BYU have built that incorporates a
lot of previous work in visualizing topic models and some of our own
ideas~\cite{gardner-2010-topic-browser}.

I have significant experience conducting and presenting research, and I have
sufficient background in machine learning and natural language processing to be
able to quickly contribute to ongoing research projects.  I would love to spend
my time as a PhD student working on machine interpretation or some subproblem
of it.  Speech recognition, speech synthesis, machine translation, parsing and
co-reference resolution are all interesting problems to me that I would enjoy
working on.  As statistical modeling is the foundation of most state-of-the-art
solutions to these problems, I also would enjoy doing research in that area.
[Add more personalized stuff to each university here]

\bibliographystyle{unsrt}
\bibliography{bib}
